\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[a4paper,draft,12pt,oneside,article,table]{memoir}

%% Geometry
\isopage[10]
\setlength{\topskip}{1.6\topskip} % for \sloppybuttom
\checkandfixthelayout
\sloppybottom

%% Typography
\usepackage{polyglossia,microtype,hyperref,amsmath,unicode-math,xcolor,natbib}
\definecolor{zen-red}{HTML}{B23333}    \definecolor{zen-orange}{HTML}{E57A33}
\definecolor{zen-yellow}{HTML}{F0DFAF} \definecolor{zen-green}{HTML}{5F7F5F}
\definecolor{zen-cyan}{HTML}{93E0E3}   \definecolor{zen-blue}{HTML}{336CB2}
\setdefaultlanguage{english} %polyglossia

\hypersetup{colorlinks=true,linkcolor=zen-red,citecolor=zen-green,urlcolor=zen-orange} % hyperref
\microtypesetup{final,verbose=silent} 
\setmainfont[Ligatures=TeX,Numbers=OldStyle]{Arno Pro} % fontspec
\setmonofont[Scale=MatchLowercase]{DejaVuSansMono} % fontspec

\unimathsetup{math-style=ISO,bold-style=ISO} % unicode-math
\setmathfont[Scale=MatchLowercase]{Cambria Math} % unicode-math


%% Titlepage
\setlength{\droptitle}{-3em}
\pretitle{\LARGE\par} \posttitle{\vskip 0.5em}
\newcommand{\supertitle}[1]{\gdef\suP{#1}}
\renewcommand{\maketitlehooka}{\ifx\suP\undefined\begin{center}\else\begin{center} {\scshape\suP}\fi}
\newcommand{\subtitle}[1]{\gdef\suB{#1}}
\renewcommand{\maketitlehookb}{\ifx\suB\undefined \end{center}\else\par {\large\scshape\suB}\par\end{center}\fi}

%% Header
\newcommand{\stunum}[1]{\gdef\stuN{#1}}
\copypagestyle{articlehead}{plain}
\makeoddhead{articlehead}{\color{gray}\theauthor\ifx\stuN\undefined\else\ifx\stuN\empty\else~(\stuN)\fi\fi}{}{\color{gray}\thedate}
\pagestyle{articlehead}

%% Grapics
\usepackage{tikz,pgfplots,tikz-timing}
\usetikzlibrary{mindmap,arrows,positioning,shapes}

%% TÅGEKAMMERET
%\usepackage{tket,tkvc}
%\newfontface\bbface[Scale=0.87]{TeX Gyre Termes Math} \TKsetup{C = {\bbface\kern-0.1exℂ}} % fontspec,tket

%% resten
\usepackage{threeparttable,siunitx,pdfpages,algpseudocode,algorithm}
\sisetup{per-mode=symbol}
\makeatletter \renewcommand{\ALG@name}{Algoritme}\makeatother
\usepackage{minted} % requires minted > 2.0-alpha2
\usemintedstyle{tango}

\newcommand{\srcpath}{../ex09/src/main/java/ddist}
\newcommand{\inmnt}[3]{\vspace{1em}\noindent\texttt{\color{gray}File: #3}\vspace{-1em}\inputminted[tabsize=4,firstline=#1,firstnumber=#1,lastline=#2,linenos]{java}{\srcpath/#3}}
\newcommand{\mil}[1]{\mintinline{java}{#1}}

%% Help
\usepackage{lipsum}
\usepackage[margin,draft]{fixme} \fxusetheme{color}

\begin{document}
\supertitle{Distributed Systems}
\title{Exercise 9}
%\subtitle{}
\author{Richard~Möhn~\small{(201311231)} \and Mathias~Dannesbo~\small{(201206106)}}
%\stunum{201311231, 201206106}
\date{\today}
\maketitle

\chapter{Introduction}
We used pairprogramming for all the code and ``pairreporting'' for the report, so we share the workload at 50\% each.

\chapter{Code overview}
We need to preform actions when the \mil{Listen}, \mil{Connect} and \mil{Disconnect} items are clicked in the menus. The items represent \mil{javax.swing.Actions} whose \mil{actionPerformed()} methods with our code.

\section{Listen}

The \mil{Listen} is parsing the server information from the \textsc{gui} and putting a new Runnable in the \textsc{awt} system \mil{EventQueue} whitch start up a new \mil{ServerSocket} for listening for another editor to connect as a client. Then it calls \mil{startCommunication} which is described in section \ref{sec:com}. Finally it set the title of the editor.

\inmnt{139}{193}{DistributedTextEditor.java}

\section{Connect}
The \mil{Connect} first clears the textareas, then parses the server information from the \textsc{gui} and connect to the corresponding socket. When the \textsc{tcp/ip} connection is etablished, it calls the \mil{startCommunication} which is described in section \ref{sec:com}. Finally it set the title of the editor.

\inmnt{195}{228}{DistributedTextEditor.java}

\section{Communication between editors}
\label{sec:com}
\tikzset{
  global/.style={draw=black,line width=1pt, inner sep=.5em, minimum height=2em, minimum width=11em, text centered},
  c/.style={global, cloud, aspect=2, cloud puffs=16,},
  n/.style={global, rectangle},
  r/.style={n, fill=red!20},
  g/.style={n, fill=green!20},
  b/.style={n, fill=blue!20},
  y/.style={n, fill=yellow!20},
  l/.style={>=latex',line width=1.5pt},
  sa/.style={l,->,shorten >=1pt},
}
\begin{figure}[hbp]
  \centering
  \caption{\mil{MyTestEvent} path through the system.}
  \begin{tikzpicture}[node distance=1cm, auto]
    \footnotesize
    \node[c] (net) {Network};
    \node[b, above=1em of net] (esen) {EventSender};
    \node[g, above=1.5em of esen] (oqueue) {Outbound queue};
    \node[r, above=1.5em of oqueue] (dec2) {DocumentEventCapturer};
    \node[b, below=1em of net] (erec) {EventReceiver};
    \node[g, below=1.5em of erec] (iqueue) {Inbound queue};
    \node[r, below=1.5em of iqueue] (erep2) {EventReplayer};
    \node[r, left=3em of dec2] (dec) {DocumentEventCapturer};
    \node[g, left=3em of net] (queue) {Queue};
    \node[r, left=3em of erep2] (erep) {EventReplayer};

    \node[above=1em of dec] (bef) {Provided};
    \node[above=1em of dec2] (aft) {Network enabled};


    \draw[sa] (dec2.south) --  (oqueue.north);
    \draw[sa] (oqueue.south) --  (esen.north);
    \draw[sa] (esen.south) --  (net.north);
    \draw[sa] (net.south) --  (erec.north);
    \draw[sa] (erec.south) --  (iqueue.north);
    \draw[sa] (iqueue.south) --  (erep2.north);

    \draw[sa] (dec.south) --  (queue.north);
    \draw[sa] (queue.south) --  (erep.north);
  \end{tikzpicture}
  \label{fig:event}
\end{figure}
In the not-networked editor the \mil{DocumentEventCapturer} and the
\mil{EventReplayer} worked together directly and communicated over one
queue: When the user wrote in the upper pane, the
\mil{DocumentEventCapturer} recorded the edit events and put them in the
queue. The \mil{EventReplayer} retrieved them from the queue and applied
them to the lower pane.

Now we have two editors (\textsc{a} and \textsc{b}) with one lower and
one upper pane on each of them. Edit events from the upper pane of
\textsc{a} have to be transmitted to the lower pane of \textsc{b} and
vice versa. We therefore need two queues: One for the transmission
from \textsc{a} to \textsc{b} and one for the transmission from
\textsc{b} to \textsc{a}.

However, since the editors are different processes on possibly
different machines, the queues have to be split up: Each has a head on
one of the peers and a tail on the other.  Not using \textsc{rmi},
\mil{EventSender} and \mil{EventReceiver} establish the connection
between the queues over the network manually. Overall it works like
this: \mil{DocumentEventCapturer} puts events in a queue as before and
\mil{EventReplayer} retrieves events from a queue like before. But
now, those are different queues.  \mil{EventSender} retrieves the
elements \mil{DocumentEventCapturer} put in an outbound queue and
sends them over the network.  \mil{EventReceiver} receives events from
the network and puts them in an inbound queue from which
\mil{EventReplayer} feeds. This establishes a persistent an
asynchronous way of communication.

Since the network operations \mil{writeObject} and \mil{readObject} can
play the role of blocking communication, queues are not absolutely
necessary. -- We could have modified the \mil{DocumentEventCapturer} so
that it sends events over the network instead of putting them in a queue
and we could have modified the \mil{EventReplayer} so that it receives
events from the network instead of taking them out of a queue. But
adding a layer (layered architecture) between those classes and the
network had the advantage that we didn't have to change them very much.
On top of that, it would have been a bad design: Classes should only be
responsible for one thing at a time. The message-queueing communication
style also prevents the \textsc{ui} thread from being blocked or slowed
down by network communication and thereby increases distribution
transparency.

Continuing to use queues enable us to leave \mil{EventReplayer} and
\mil{DocumentEventCapturer} largely unchanged. We just had to pull out
the queue the latter had contained before and make the former take
events directly from a queue instead of indirectly through the
\mil{DocumentEventCapturer}. The following listings show the results of
these changes.

\inmnt{1}{1000}{EventReplayer.java}

\inmnt{43}{67}{DistributedTextEditor.java}

\inmnt{30}{39}{DocumentEventCapturer.java}

To set up the communication, the \mil{Listen} and \mil{Connect} actions
call the method \mil{startCommunication} of the
\mil{DistributedTextEditor} shown below. It starts the \mil{EventSender}
and \mil{EventReceiver} as new threads as they have to send and receive
events asynchronously.

\inmnt{300}{318}{DistributedTextEditor.java}

The following two listings show the \mil{EventSender} and
\mil{EventReceiver} which connect queues over the network as described
above.

\inmnt{1}{1000}{EventSender.java}

\inmnt{1}{1000}{EventReceiver.java}

\section{Disconnect}

\mil{Disconnect} is a menu item like \mil{Listen} and \mil{Connect} and
after it is clicked, the corresponding \mil{actionPerformed} method
printed below resets the editor to a disconnected state and kicks off
the teardown of the connection.

\inmnt{230}{240}{DistributedTextEditor.java}

Surprisingly, connection teardown is much more difficult to implement
than connection setup, because a number of threads have to be notified
that they should end their life. Since the only way to reach all threads
is the events we already use for communicating text edit actions, we
introduce a special event, the \mil{DisconnectEvent}. By putting it in
the outbound queue, the \mil{Disconnect} menu action triggers a rather
complicated process of closing the connection. The following listing
contains the code of the \mil{DisconnectEvent} along with a description
of process of closing the connection.

\inmnt{1}{1000}{DisconnectEvent.java}

%\inmnt{1}{10}{MyTextEvent.java}
%\inmnt{1}{10}{TextInsertEvent.java}
%\inmnt{1}{10}{TextRemoveEvent.java}

\chapter{Conclusion}
\fxerror{List of concepts}

We have changed the provided rudimental text editor so that instances of
it are capable of sending editing events over a network to each other
and replaying them locally. The underlying means of communication is
TCP, since it offers reliable and ordered transmission of data. Our
application comprises two layers on top of the already layered TCP/IP
stack and establishes persistent asynchronous communication by means of
message queueing.

The hardest part in developing the system was to figure out how four
threads and a networking connection should be torn down in a clean and
ordered way. It is also here where some unexpected and yet unexplained
behaviour still occurs.

There are some more points that might be improved: When an editor
listens for connections, it does so in the main thread, so that the
\textsc{gui} freezes. The \mil{DisconnectEvent} extending
\mil{MyTextEvent} is a case of implementation inheritance which isn't
appropriate in this case. Instead, both should implement a common
\mil{Event} interface. Instead of handling exceptions, the editor
terminates. Classes are tagged Serializabl without caring for the
implications of this.

All in all, however, we came up with a rather elegant design, which
should make clearing up the above issues easy.

%\clearpage
% \listoftables
% \listoffigures
% \listoflistings
%\nocite{*}
%\bibliographystyle{dlfltxbbibtex} \Bibliography{bib}
%\clearpage \appendix

\end{document}

%%% Local Variables:
%%% coding: utf-8
%%% mode: latex
%%% TeX-engine: xetex
%%% End:
